Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-11 article/article.tex
yes

Table~\ref{tab:results} illustrates the overall performance of the three classifiers. Rows indicate various feature selection methods. We applied PCA to the original data, while the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}). For instance, \emph{Norm/Vox-M ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation. Columns depict the used classifiers. For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.

@@

From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used. By using an exact binomial test we found that GNB that used PCA as feature extraction method performed significantly better than chance-level, with a 95\% $CI: 56.9 - 65.8$, $p < 0.001$. However, a Chi-square test for equality of proportions indicated that using PCA as dimentionality reduction method performed significantly worse than a GNB classifier on the means of the selected ROI's, $\chi^2(1) = 7.7848, p < 0.01$. This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information. Additionally, the data may not suited to be aligned by voxel index (cf.~Section~\ref{sec:dimensionalityReduction}). The performance listed in Table~\ref{tab:results} is for PCA using 149 components. This value was found by performing classification with a varying amount of components, the results for which are listed in Figure~\ref{fig:pca_components}.

>>>>

Table~\ref{tab:results} illustrates the overall performance of the three classifiers. Rows indicate various feature selection methods. We applied PCA to the original data, while for the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}). For instance, \emph{Norm/Vox-M ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation. Columns depict the used classifiers. For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.

@@

From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used. By using an exact binomial test we found that GNB that used PCA as feature extraction method performed significantly better than chance-level, with a 95\% $CI: 56.9 - 65.8$, $p < 0.001$. However, a Chi-square test for equality of proportions indicated that using PCA as dimentionality reduction method performed significantly worse than a GNB classifier on the means of the selected ROI's, $\chi^2(1) = 7.7848, p < 0.01$. This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information. Additionally, the data may not be well suited to be aligned by voxel index (cf.~Section~\ref{sec:dimensionalityReduction}). The performance listed in Table~\ref{tab:results} is for PCA using 149 components. This value was found by performing classification with a varying amount of components, the results for which are listed in Figure~\ref{fig:pca_components}.
